---
title: "<center><div class='mytitle'>Principal Component Analysis</div></center>"
output:
  html_document:
      highlight: arrow
      css: style.css
      toc: FALSE
      #math_method: 
        #engine: webtex
        #url: https://latex.codecogs.com/svg.image?
      citation_package: biblatex
      #includes: 
         #in_header: Capa.html
         #after_body: Footer.html
header-includes:
  - \usepackage{amsmath}
bibliography: yourBibFile.bib
link-citations: true
biblio-style: authoryear
---



```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(tidy.opts=list( width.cutoff = 30 ), tidy=T)

```

<section class="section" >
  <div class="content">

In this topic, I will discuss the mathematics of  Eigenvectors and Eigenvalues, Spectral Decomposition Theorem, and Principal Component Analysis. And finally, I will show an application of this technique. 

# Eigenvectors and Eigenvalues

<div class="Definition">
<p><strong>Definition 1</strong>.
Let a non-null vector $v$ (i.e. $\vec{v} \neq 0$), 
</p>
</div>

The first one is the most famous distribution, is the Bernoulli distribution. Let $X$ a binary random variable with probability density function (PDF) $f_{x}$ . Then, $X \sim Ber(p)$ has PDF

<div class="equation">
  $$f(x) = \P(X = x) = p^{x} (1-p)^{1-x}$$
</div>
    
where the support is $X \in \{0, 1\}$ and parametric space is $p \in (0, 1)$. The expected value (mathematical expectation) is $\E(X) = p$ and variance is $\var(X) = p(1 - p)$.

I will not discuss moments in statistics here where the first moment is mathematical expectations and the second is related to variance. Nevertheless, [Wikipedia](https://en.wikipedia.org/wiki/Moment_(mathematics)) is a good site where you might start to study more about this topic. I love this concept because everything concerning statistical models is linked to a mean, mainly in generalized linear models ([MLG](https://en.wikipedia.org/wiki/Generalized_linear_model)). But it is a topic to see forward. 


<span style="width:40px;display:inline-block"></span>

```{r warning=FALSE, message=FALSE, fig.align="center", fig.height=4, class.source="watch-out"}
set.seed(123)

```



 
# REFERENCES

<div id="refs">
---
nocite: '@*'
---
</div>

</div>
</section>





